---
title: "p8105_hm1_xl3371"
author: "Xuer Liu"
date: "2023-09-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse) #load packages
library(moderndive)
data("early_january_weather")
```

# Problem 1

## Description of the dataset
The dataset "early_january_weather" contains 15 variables which are `r colnames(early_january_weather)`. 

Some importants variables are "temp"and "humid". 

For variable "temp", the mean value is `r mean(early_january_weather$temp)` and standard deviation (SD) is `r sd(early_january_weather$temp)`. 

For variable "humid", the mean value is `r mean(early_january_weather$humid)` and SD is `r sd(early_january_weather$humid)`.

The dataset consists of `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns.

The mean temperature is `r mean(early_january_weather$temp)` degrees.

## Create a scatterplot
```{r,message=FALSE}
plot <- early_january_weather %>% ggplot(aes(x=time_hour, y=temp, color=humid)) +
  geom_point(alpha=0.5) +
  labs(x="Time Hour",
       y="Temperature",
       color="Humidity")
print(plot)
ggsave("plot.png")
```
Above is the scatterplot which represents the relationship between temperature and time in hour for different levels of humidity. The plot shows there is a strong positive and linear association between the two variables, temperature and time hour, under the levels of humidity with some potential outliers. As the temperature and time increase, the humidity level increases.

# Problem 2

## Create a data frame
```{r}
df <- tibble(
  numeric_vector = rnorm(10),
  logic_vector = numeric_vector > 0,
  character_vector = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  factor_vector = factor(c("medium", "mild", "hot", "medium", "mild", "hot", "medium", "mild", "hot", "medium"))
)
```

## Take the mean of each variable
```{r}
mean(df %>% pull(numeric_vector)) #work
mean(df %>% pull(logic_vector)) #work
mean(df %>% pull(character_vector)) #return NA
mean(df %>% pull(factor_vector)) #return NA
```
We could successfully take the mean of numeric and logical variables; however, we are not able to take the mean of character and factor variables.

## Convert variables
```{r, eval=FALSE}
as.numeric(df %>% pull(logic_vector))
as.numeric(df %>% pull(character_vector))
as.numeric(df %>% pull(factor_vector))
```
When converting the logical vector to numeric vector, the result becomes 1's and 0's respectively.The logical vector output contains TRUE and FALSE, As seen in the output, the TRUE values in the logical vector have been converted to 1, and the FALSE values have been converted to 0. This also explain why we can get the mean of logical vector.

When converting the factor vector to numeric vector, the output is 1, 2,and 3, since we only have 3 different factor "levels". Rstudio automatically assign different factor levels with integer codes, and the mean of the integer codes does not equate to the mean of the original data. Thus, we are not able to take the mean of factor vector.

As for character vector, we are not able to convert it to a numeric vector, it returns NA since there is no numeric value, which explains why we cannot take the mean of character vector since it is meaningless to calculate the mean of non-numeric stings.

